Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.

### 2023-09-08 inst/manuscript/revision_2/manuscript_microxanox_rev2.Rmd
Reorganisation of folder
Reorganisation step 2 and recompilation of rev 2

---
title: "Microxanox: an R package for simulating an aquatic $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch;  Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
keywords: 
  - reproducibility
  - regime shift 
  - final state
  - ordinary differential equations
abstract: |
  $Microxanox$ is an R package to simulate a three functional group ecosystem (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations. Simulations can be run individually or over a parameter range.  The model can be implemented with different numbers of species per functional group. The package is constructed in such a way that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated. Furthermore, the package framework and code should serve as a useful starting point for making simulation models of other types of ecosystems.
journal: "SoftwareX"
date: "`r Sys.Date()`"

@@

# Citation settings
bibliography: bibliography.bib
csl: 'softwarex.csl'
nocite: '@Bush2017'
linkcolor: blue
link-citations: yes
---

@@

```{r uc1, echo=FALSE, fig.align="center", fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the final states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\\_dynamics()$  function.', message=FALSE, warning=FALSE, include=TRUE, messages=FALSE, out.width=350}
fig <- here::here("inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching.pdf")
if (!file.exists(fig)){
  library(ggplot2)
  library(microxanox)
  library(dplyr)
  x <- readRDS(here::here("inst/manuscript/data/hlh_with_defaults.RDS"))
  p <- plot_dynamics(x)
  p[[1]] <- p[[1]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[2]] <- p[[2]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[3]] <- p[[3]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p_a <- p[[4]]
  p[[4]]$data <- p[[4]]$data %>% dplyr::filter(species != "a")
  p[[4]]$data$species <- p[[4]]$data$species %>%
    dplyr::recode(SO = "Oxidised sulfur", O = "Oxygen", SR = "Sulfide", P = "Phosphorus")
  p[[4]] <- p[[4]] +
    xlab("time (h)") + 
    ylab(expression(atop(log[10](concentration),(µM)))) + 
    labs(colour = "substrate") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p_a$data <- p_a$data %>% dplyr::filter(species == "a")
  p_a$data$species <- p_a$data$species %>%
    dplyr::recode(a = "Oxygen diffusivity ")
  p_a <- p_a +
    ylab(expression(atop(Oxygen~diffusivity,(hr^{-1})))) + 
    labs(colour = "Environmental driver") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p <- p[[1]] / p[[2]] / p[[3]] / p[[4]] / p_a
  ggsave(
    filename = fig,
    plot = p,
    device = "pdf",
    width = 8,
    height = 8
  )  
}
knitr::include_graphics(fig)
```

@@

```{r uc2, echo = FALSE,  include=TRUE, fig.cap='The amount of hysteresis depends on the combination of types of organisms present. The model is entirely deterministic, hence there are no error bars.', fig.align="center", out.width = 150 }
fig <- here::here("inst/manuscript/figures/gen_uc2_user_guide_hysteresis.pdf")
if (!file.exists(fig)){
  library(ggplot2)
  x <- readRDS(here::here("inst/manuscript/data/resp_summ5.rds"))
  x$composition <- as.character(x$composition)
  x <- x[order(x$mean_mean_hyst),]
  x$composition <- factor(x$composition, x$composition, ordered = TRUE)

>>>>

---
title: "Microxanox: an R package for simulating an aquatic $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch;  Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
keywords: 
  - reproducibility
  - regime shift 
  - final state
  - ordinary differential equations
abstract: |
  $Microxanox$ is an R package to simulate a three functional group ecosystem (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations. Simulations can be run individually or over a parameter range.  The model can be implemented with different numbers of species per functional group. The package is constructed in such a way that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated. Furthermore, the package framework and code should serve as a useful starting point for making simulation models of other types of ecosystems.
journal: SoftwareX
date: "`r Sys.Date()`"

@@

# Citation settings
bibliography: bibliography.bib
csl: './../softwarex.csl'
nocite: '@Bush2017'
linkcolor: blue
link-citations: true
---

@@

```{r uc1, echo=FALSE, fig.align="center", fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the final states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\\_dynamics()$  function.', message=FALSE, warning=FALSE, include=TRUE, messages=FALSE, out.width=350}
fig <- "./figures/gen_uc1_partrep_temporal_state_switching.pdf"
if (!file.exists(fig)){
  library(ggplot2)
  library(microxanox)
  library(dplyr)
  x <- readRDS(here::here("inst/manuscript/data/hlh_with_defaults.RDS"))
  p <- plot_dynamics(x)
  p[[1]] <- p[[1]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[2]] <- p[[2]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[3]] <- p[[3]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p_a <- p[[4]]
  p[[4]]$data <- p[[4]]$data %>% dplyr::filter(species != "a")
  p[[4]]$data$species <- p[[4]]$data$species %>%
    dplyr::recode(SO = "Oxidised sulfur", O = "Oxygen", SR = "Sulfide", P = "Phosphorus")
  p[[4]] <- p[[4]] +
    xlab("time (h)") + 
    ylab(expression(atop(log[10](concentration),(µM)))) + 
    labs(colour = "substrate") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p_a$data <- p_a$data %>% dplyr::filter(species == "a")
  p_a$data$species <- p_a$data$species %>%
    dplyr::recode(a = "Oxygen diffusivity ")
  p_a <- p_a +
    ylab(expression(atop(Oxygen~diffusivity,(hr^{-1})))) + 
    labs(colour = "Environmental driver") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p <- p[[1]] / p[[2]] / p[[3]] / p[[4]] / p_a
  ggsave(
    filename = fig,
    plot = p,
    device = "pdf",
    width = 8,
    height = 8
  )  
}
knitr::include_graphics(fig)
```

@@

```{r uc2, echo = FALSE,  include=TRUE, fig.cap='The amount of hysteresis depends on the combination of types of organisms present. The model is entirely deterministic, hence there are no error bars.', fig.align="center", out.width = 150 }
fig <- "./figures/gen_uc2_user_guide_hysteresis.pdf"
if (!file.exists(fig)){
  library(ggplot2)
  x <- readRDS(here::here("inst/manuscript/data/resp_summ5.rds"))
  x$composition <- as.character(x$composition)
  x <- x[order(x$mean_mean_hyst),]
  x$composition <- factor(x$composition, x$composition, ordered = TRUE)

### 2023-09-08 inst/manuscript/revision_2/manuscript_microxanox_rev2.tex
Reorganisation step 2 and recompilation of rev 2

{\centering \includegraphics[width=350px]{../figures/gen_uc1_partrep_temporal_state_switching}

@@

{\centering \includegraphics[width=150px]{../figures/gen_uc2_user_guide_hysteresis}

>>>>

{\centering \includegraphics[width=350px]{./figures/gen_uc1_partrep_temporal_state_switching}

@@

{\centering \includegraphics[width=150px]{./figures/gen_uc2_user_guide_hysteresis}
